Windowed Developer Topic Analysis

### 2009-04-03 lda-paper.tex
ric edits

  At different times, developers focus on different topics and tasks
  during development.  These different topics and times of focus might
  be obvious to developers familar with the project, but other
  stakeholders, like managers, might be less aware of the topics of
  development. Imagine the scenario where developers had to deal with
  issues orthogonal to the features suggested, at the end of the
  iteration a manager would wonder where all the time went, the
  developers might have just blocked all sides issues in with features
  they were initially working on, they might not have noticed their
  focus had changed.  Currently it is difficult for stakeholders to
  audit what actually occurred and when certain features or topics
  were focussed on. We propose and provide a visualization of the
  changing development topics over time. Our approach takes a topic
  analysis tool like Latent Dirichlet Allocation (LDA) or Latent
  Semantic Indexing (LSI) and instead of applying the tool to the
  entire corpus we apply to windows of the corpus. This allows us to
  see the unique topics of development occuring during a particular
  time period and then see if these individual topics ever reappear.

@@

What happened in the previous iteration? What were the developers
working on? What was the developer in the next cubicle working on?
What were the common topics and issues dealt with in the current
release of the software? What were the topics of the previous release?
Given the requirements agreed to at the start of the iteration did our
developers work on them? What else did they work on. Were there issues
which were orthogonal to the

Topic analysis, with respect to Software Control Systems (SCSs), is
valuable as it has many uses for the stakeholders involved.  Example
scenarios where the usage of these techniques would prove useful are
numerous. Imagine a development team has agreed to implement some
features at the start of an iteration. By the end of the iteration
only some of the agreed upon features are implemented. The developers
swear they were focusing solely on the features agreed upon. Topic
analysis could help by teasing out the distinct topics or focusses of
the developers. Perhaps the developers had to deal with issues
orthogonal to the features they were working on, but were probably
necessary to complete such features. Developers might have grouped
those orthogonal changes into one of the features they were working
on, thus when asked about how their time was spent the orthogonal
changes would not be covered. This is what we hope that windowed topic
analysis can uncover. If windowed topic analysis can discover
important but overlooked topics, it can help both developers and
managers. Developer can reflect and see a good summary of what worked
on, and managers can see the focus of development while still
remaining relatively high level.

We propose to extend common concept analysis techniques by executing
them over windows of documents over time. We hope by looking at topic
clusters over time we can identify continuous effort on a select
group of topics.

Topic analysis of fine-grained revisions is important as often
revisions belonging to different topics are interleaved or even mixed
inside of commits.  Sometimes a single revision to a file contains
multiple bug fixes.  Using powerful topic analysis techniques like
Latent Semantic Indexing (LSI) or Latent Dirichlet Allocation (LDA)
one can seperate these mixed signals into coherent topics. In this
case a topic is a set of tokens that are related to each other.

@@

Another use of topic analysis is to help us partition a project's
time-line into sub-iterations of developer focus. By breaking apart an
iteration we might be able to extract the underlying software
development process from these artifacts or at the very least what
maintenance was occurring in the repository at that time.

In this paper we seek to explore how topics shift over time in the
source control repository of a project. We analyze commit comments per
revision in each time window and we extract the topics in that time
window. We expect that topics in each window will change over time and
the similarities and differences in these sets of topics will indicate
both developer focus and changes in developer focus.

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topic and trend. Messages are block of
text written by developers, in this paper messages will be the CVS
commit comments. Word distributions are the summary of a message by
word count, effectively the word distribution will be over all
messages looked at but since most words will not appear in a message,
a word distribution is effectively a word count divided by the message
size. A topic is a unique word distribution, but effectively a set of
words unique to an independant behaviour in the text corpus, in this
paper we often summarize topics by their top 10 more frequent words.
A trend is a one or more similar topics which reoccur over
time. Trends are important because they are sets of topics which
reoccur over the development of a project.

@@

Our blackbox view of LDA and LSI is that they extract topic clusters
from the documents that are independant from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familar with the
corpus. Technically we could swap LDA for LSI. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

@@

@@

\section{Appendix}

\begin{itemize}
\item We are missing a comparison of our technique to the classical technique
\end{itemize}

>>>>

  At different times, developers focus on different topics and tasks
  during development.  These different topics and tasks might
  be obvious to developers familar with the project, but other
  stakeholders, like managers, might be less aware of the topics of
  development. Imagine the scenario where developers unexpectedly had to deal with
  issues orthogonal to the features suggested, at the end of the
  iteration a manager would wonder where all the effort went to.
  Currently it is difficult for stakeholders to
  audit what actually occurred and when certain features or topics
  were focussed on. We propose a visualization of the
  changing development topics over time. Our approach uses a topic
  analysis tool like Latent Dirichlet Allocation (LDA) or Latent
  Semantic Indexing (LSI). Instead of applying the tool to the
  entire corpus we apply it to windows (periods) of the corpus. This allows us to
  see the evolving stream of topics of development occuring
  over time.

@@

What happened in the previous maintenance or development iteration? What were the developers
working on? What was the developer in the next cubicle working on?
What were the common topics and issues dealt with in the current
release of the software? What were the topics of the previous release?
Given the requirements agreed to at the start of the iteration did our
developers work on them? What else did they work on.

Topic analysis, with respect to Software Control Systems (SCSs) like
CVS, Subversion, Git etc., is potentially valuable as it has many uses
for various stakeholders and uses.
Imagine a
development team has agreed to implement some features at the start of
an iteration. By the end of the iteration only some of the agreed upon
features are implemented. The developers swear they were focusing
solely on the features agreed upon. Topic analysis could help by
teasing out the distinct topics or focusses of the developers. Perhaps
the developers had to deal with issues.
Developers might have grouped those orthogonal changes into
one of the features they were working on, thus when asked about how
their time was spent the orthogonal changes would not be covered. This
is what we hope that windowed topic analysis can uncover. If windowed
topic analysis can discover important but overlooked topics, it can
help both developers and managers. Developer can see a
summary of what worked on, and managers can development.

Topic analysis of revisions in SCSs is important as often revisions
belonging to different topics are interleaved or even mixed inside of
commits.  Sometimes a single revision to a file contains multiple bug
fixes.  Using powerful topic analysis techniques like Latent Semantic
Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
these mixed comments into topics. These topics describe seperate ideas
that are interleaved throughtout the development. They indicate groups
of comments which are related to each other by their language, often
the content of the comments are related.  In this case a topic is a
set of tokens that are related to each other.

@@

One use of topic analysis is to help us partition a project's
time-line into periods  of developer focus. By breaking apart an
iteration we may be able to recognize the underlying software
development process from these revision or at the very least what
maintenance was occurring in the repository at a particular time.

We propose to extend the use of common topic analysis techniques by executing
them over windows of documents over time. We hope by looking at topic
clusters over time we can identify continuous effort on a select
group of topics.

In this paper we explore how topics shift over time in the
source control repository of a project. We analyze commit comments per
revision in each time window and we extract the topics in that time
window. We expect that topics in each window will change over time and
the similarities and differences in these sets of topics will indicate
both developer focus and changes in developer focus.

Our contributions include:
\begin{itemize}
\item Demonstrating the value of windowed topic analysis
\item Multiple visualizations of trends over time
\end{itemize}

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topic and trend. \emph{Messages} are
blocks of text written by developers. In this paper messages will be
the CVS commit comments made when the user commits changes to files in
a repository. Word distributions are the summary of a message by word
count. The word distribution will be over all messages
looked at but since most words will not appear in a message, a word
distribution is effectively a word count divided by the message
size. A topic is a unique word distribution, but effectively a set of
words unique to an independant behaviour in the text corpus, in this
paper we often summarize topics by their top 10 more frequent words.
A trend is a one or more similar topics which reoccur over
time. Trends are important because they are sets of topics which
reoccur over the development of a project.

@@

Each of these, LDA and LSI, extract topic clusters
from the documents that are independent from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familar with the
corpus. Technically we could swap LDA for LSI. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

@@

Even with our liberal topic similarity metrics which produced both
long and short trends, we showed that there are only a few trends in a
repository which re-appear. Since so few trends re-appear and so many
trends appear only once this suggests that global topic analysis might
be ignoring locally unique topics.

The utility of global topic analysis is questionable if the value of
information decreases as it becomes older. Perhaps older trends will
dominate the topic analysis. Windowed localized topic analysis show
what are the unique topics.

We presented multiple visualizations which focused on different
aspects of the data: temporality of trends, trend size, and a
condensed overall trend view. The condensed trend view shows much more
information than the views that global analysis could show.

@@

% \section{Appendix}

% \begin{itemize}
% \item We are missing a comparison of our technique to the classical technique
% \end{itemize}
